Examining a Quarter Century of Publishing Trends in Social Work Research:

### 2015-05-18 manuscript.Rmd
Updated figures.

x <- c(1:10)
y <- c(1:10)

df <- data.frame(x,y)
pg <- ggplotGrob(plotSmallMultiples.1)

pl <- gtable_filter(pg, 'panel', trim=F)$layout
pg <- gtable_add_grob(pg, x.grob, t=max(pl$t), l=max(pl$l))

grid.newpage()
grid.draw(pg)

```{r uncitedSummary}
#=======================================
# Create new data file 
#=======================================

articles.df <- left_join(n.so, n.yr) %>%
    group_by(title, year) %>%
    summarize(N = n()) %>%
    mutate(Overall.N = sum(N), year = as.numeric(year)) %>%
    ungroup() %>%
    arrange(desc(Overall.N), title) %>%
    rename(Year = year, Journal= title)

scopus.df <- read.csv("/Users/beperron/Git/SocialWorkResearch/Data/scopus.csv")
articles.df <-full_join(sm.df, scopus.df, by = c("Journal", "Year"))

articles.df<- articles.df %>% 
    filter(Year <= 2013) %>%
    mutate(propNotCited = X..Not.Cited/100) %>%
    select(-X, -SJR, -SNIP, -Citations, -IPP, -X..Not.Cited, -Overall.N, -X..Review) %>%
    rename(HD.N = N, scopus.N = Documents)

#=======================================
# Calculate number of uncited articles
#=======================================

#  First reduce the data file to 1996 to 2010. Data contained significant problems that brought the number of uncited values to almost 0.  Was not reliable.

df.96.10 <- filter(articles.df, Year >= 1996, Year <= 2010)

# If HD article count is missing, replace with Scopus count

# Some counts have 0 values.  Set these to NA and then omit.
df.96.10$HD.N <-ifelse(df.96.10$HD.N == 0, NA, df.96.10$HD.N) 
df.96.10 <- df.96.10[!is.na(df.96.10$HD.N), ]

# If uncited proportion is missing, compute average for that particular journal
# Replace missing value with the mean value
df.96.10 <- df.96.10 %>% group_by(Journal) %>%
        mutate(meanPropUncited = mean(propNotCited, na.rm=TRUE)) %>% ungroup()

df.96.10$propNotCited <- ifelse(is.na(df.96.10$propNotCited),
     df.96.10$meanPropUncited, df.96.10$propNotCited)

# If uncited proportion is missing because it was not in Scopus, replace the missing
# value with the overall mean uncited proportion

overallPropUncited <- mean(df.96.10$propNotCited, na.rm=TRUE)

df.96.10$propNotCited <- ifelse(is.na(df.96.10$propNotCited), overallPropUncited, df.96.10$propNotCited)

# Calculate number of not cited

df.96.10 <- df.96.10 %>% mutate(countNotCited = HD.N * propNotCited)

df.96.10.Summary <- df.96.10 %>% group_by(Year) %>%
    summarize(Published = sum(HD.N), 
              Uncited = sum(countNotCited))

#============================================
# Create a melted data frame for the plot
#============================================

# We need to include years before and after the scopus data.

df.95.11 <- select(articles.df, Year, HD.N) %>% 
    filter(Year <= 1995 |Year >=2011) %>%
    rename(published = HD.N) %>%
    group_by(Year) %>%
    summarize(Published = sum(published, na.rm=TRUE)) %>%
    ungroup() %>%
    mutate(Uncited = NA)

uncited.df <- rbind(df.96.10.Summary, df.95.11) %>%
    arrange(Year)

#============================================
# Compute percentage variable and melt
#============================================

uncited.df$percentage <- uncited.df$Uncited/uncited.df$Published*100

uncited.melted <- reshape2::melt(uncited.df[,c("Year", "Published", "Uncited")], 
     id.vars = c("Year"))

#=======================================
# Create text annotations to plot 
#=======================================

percent.df <- uncited.df
percent.df$percentage <- round(percent.df$percentage, digits=1) 
percent.df$percentLabel<- paste(percent.df$percentage, "%", sep="")
percent.df <- na.omit(percent.df)
percent.df <- percent.df[c(seq(1, 15, 3), 15),]

#=======================================
# Create the plot 
#=======================================

cp <- c("#5e3c99", "#e66101")

plotArticles <- ggplot(uncited.melted, aes(Year, value, colour = variable,)) + 
    geom_line(alpha = .60) + 
    geom_point(size = 4) +
    scale_color_manual(values=cp, labels = c("Number of Articles Published",
         "Estimated Number of Uncited Articles")) +
    theme(legend.title=element_blank()) +
    theme(legend.position = "bottom") +
    ylab("N") +
    geom_text(data = percent.df, aes(Year, Uncited -70, label = percentLabel), size =4, colour = "black") +
    scale_y_continuous(breaks = seq(0, 2250, 250), limits = c(0, 2250))

```{r}
# Read Scopus database
scopus.df <- read.csv("/Users/beperron/Git/SocialWorkResearch/Data/scopus.csv")

boxplot.scopus <- ggplot(subset(scopus.df, subset =Year <=2013 & Year >= 1999), aes(as.factor(Year), IPP)) +
    geom_boxplot()

```{r plotJournalCount, eval=TRUE}

@@

plot.journal.count <- ggplot(n.journals.year, aes(as.factor(record.y), 
    y = Journal.count, group=1)) + 
    geom_line(colour="black") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=8), 
    title = element_text(size = 10), plot.title=element_text(size=10)) + 
    xlab("Year") + 
    ylab("Journals") + 
    ggtitle("B. Number of Journals by Year") +
    scale_x_discrete(breaks=c(seq(1989, 2013, 2))) +
    scale_y_continuous(breaks = c(seq(0, 100, 10)), limits = c(20, 70)) 
```

@@

plot.author.count <- ggplot(n_authors2, aes(x = year, group=1)) + 
    geom_line(aes(y=Median, colour="Median"), size = .75, colour = "#fdb864", alpha = .70) + 
    geom_line(aes(y=Average, colour = "Average"), size = .75,  colour = "#e66101" ) +
    geom_point(aes(y=Average, size = Standard_Deviation), colour = "#e66101") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), 
    title = element_text(size = 10)) +  
    scale_x_discrete(breaks=c(seq(1989, 2013, 2))) +
    scale_y_continuous(limits = c(0,2.5)) +
    ylab("Value") +
    xlab("Year") +
    ggtitle("Authors Per Article by Year") +
    labs(size = "Standard deviation") 
```

@@

journal.evidence.plot <- ggplot(na.omit(df1), aes(x = N, y= journal, group = journal, colour = studyType)) + 
    geom_segment(aes(yend = journal), xend=0, colour="grey50") +
    geom_point(size = 2.5) + 
    scale_color_manual(values = cp) + 
    theme_bw() +
    facet_grid(.~studyType) + 
    theme(panel.grid.major.y = element_blank(), legend.position = "none") + 
    xlim(0, 600) + 
    ylab("Journal") +  
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    scale_x_continuous(breaks = seq(0, 60, 10))

@@

@@

page.plot <- ggplot(pages.df, aes(years, Mean.Pages, size = Std.Pages, group = 1))+ 
    geom_line(size=.75, colour =  "#5e3c99") +
    geom_line(aes(years, Median.Pages, group = 1), size = .75, colour = "#b2abd2", alpha = .70) + 
    geom_point(colour = "#5e3c99") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), 
    title = element_text(size = 10)) +
    scale_x_discrete(breaks=c(seq(1989, 2013, 2))) +
    scale_y_continuous(limits = c(11,16)) +
    ylab("Value") +
    xlab("Year") +
    labs(size = "Standard deviation", colour = "Median")  +
    ggtitle("Article Length by Year")

@@

# Results

@@

A total of XX article records from XX unique journal titles met the inclusionary
criteria and were included in the historical database.

Figure 1 uses a
small multiples plot to summarize the number of article records for each year
throughout the study period.  Small multiples, also referred to as trellis
and lattice charts, are the best design solution for visually enforcing comparisons of
change (Tufte, XXXX).  In this context, we use two different data sources to show differences in publishing across the journals -- the orange line represents data from the SWHD, and the purple line represents data from Scopus.

the small multiples to show differences in publishing across the journals,

show differences in publication history for each journal.

The number of article records for each year (by journal) is denoted by an orange line, along with the overall number of article records.  For further contextual information, the number of article records available from Scopus is presented with a purple line.   
This plot reveals variability in publishing over time and across journals.  The for each journal and across journals.  
First, the historical database indexed XX journals, whereas Scopus indexed only XX.  For the number of journals indexed by Scopus, the number of article records tend to be higher than the number of records available in the historical database.  The primary reason for this difference is that Scopus included editorials and book reviews in their summary data, whereas these were excluded from the SWHD.  Other difference can be attributed to errors and gaps in indexing across both databases.  As we relied upon aggregate data from Scopus, we were unable to quantify these differences.

should be noted that the number of article records available from Scopus tends to be greater than the historical database because Scopus included editorials (AND REVIEWS), whereas these were excluded from the historical database.

According to Tufte (Evisioning information), small multiples are the best design s

Table 1 provides a summary of the journal titles and corresponding article records collected from the databases.  This table includes the specific years each journal is represented in the current study, the number of corresponding article records, the journal's h-Index, and whether the journal was included on the list of social work disciplinary journals defined by Hodge and Lacasse (2011). Appendix A provides a summary of journals that were on the list by compiled by Hodge and Lacasse (2011) but not included in this study (N = `r length(hodge.swHistory.diff)`), in addition to the journals excluded from the study due to an article count $<=$ 20 (N = `r nrow(fewArticles)`).  It is important to note that the years of inclusion noted for each journals does not correspond to its full historical record.  Many journals have publications that both precede and proceed the 25-year timeframe of the current study.  Moreover, many journals are missing articles records due to delays and errors in indexing, which is a known and longstanding problem in social work research (HoldenXXXX). These issues are given further attention in the discussion section of the article.

```{r}
# As shown in Table 1, `r nrow(final.list.25)` journals were included in the current study.  The mean number of years of inclusion is `r options(digits=2); mean(final.list.updated$study.time)` (median = `r options(digits=2); median(final.list.updated$study.time);`, sd =  `r options(digits=2); sd(final.list.updated$study.time)`). The overall h-Index for each journal.  The h-Index values represent the current h-Index of each journal, as opposed to the h-Index of the final study period, so these values should be interpreted cautiously.  The mean h-Index value is XXXX, (median = XXX; sd = XXX). A total of `r pretty(sum(articles.year.df$year.count))` article records were obtained across these journals.  The average number of articles per journal was `r options(digits=2); pretty(mean(articles.year.df$year.count))` (median = `r options(digits=2); pretty(median(articles.year.df$year.count))`, sd = `r options(digits=2); pretty(sd(articles.year.df$year.count))`). This figure shows consistent growth in the number of articles from 1989 to 2005.  Notable fluctuations in the number of articles was observed from 2005 to 2008, followed by growth until 2012. The decline from 2012 to 2013 should be interpreted cautiously, as this may be an artifact due to delays in indexing.  
```
Data from the National Science Foundation regarding the number of earned doctorates in social work and social welfare were added to further contextualize the publication trends.  These data reveal both growth and decline in the number of earned doctorates over the period of time for which data were available.  The number of earned doctorates exhibited a rise from 2003 to 2005, which corresponded to a rise in the number of article published during these years.  Between 2008 and 2012, the number of doctoral degrees exhibited a decline, but the number of published articles continued to rise.

@@

```
\newpage

@@

```{r}
plot.journal.count
```

@@

```{r}
stacked.plot
```
\newpage
```{r fig.height=9.5}
journal.evidence.plot
```

\newpage
```{r fig.height=9.5}
EBP.plot
```

>>>>

# x1.grob <- ggplotGrob(plotSmallMultiples.1)
# x2.grob <- ggplotGrob(ggplot(cars, aes(speed, dist)) + geom_point())
# 
# library(gtable)
# plot1 <- gtable_filter(x1.grob, 'panel', trim=F)$layout
# pg <- gtable_add_grob(x1.grob, x2.grob, t = max(plot1$t), l = max(plot1$l))
# grid.newpage()
# grid.draw(pg)
```

```{r articleSummary}
#=======================================
# Create new data file 
#=======================================
library(gridExtra)

articles.df <- left_join(n.so, n.yr) %>%
    group_by(year) %>%
    summarize(N = n()) %>%
    mutate(year = as.numeric(year), CumSum = cumsum(N)) %>%
    ungroup() %>%
    arrange(desc(N), year) %>%
    rename(Year = year)

plot.article.count <- ggplot(articles.df, aes(Year, N, group =1)) + 
    geom_line(colour=cp[2], size = 1) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=8), 
    title = element_text(size = 10), plot.title=element_text(size=10)) + 
    xlab("Year") + 
    ylab("") + 
    ggtitle("Article Count") +
     scale_x_continuous(breaks=seq(1989, 2015, 8)) +
        scale_y_continuous(breaks = c(seq(0, 2200, 250)), limits= c(500, 2100))

plot.article.cumulative <- ggplot(articles.df, aes(Year, CumSum, group =1)) + 
    geom_line(colour=cp[2], size = 1) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=8), 
    title = element_text(size = 10), plot.title=element_text(size=10)) + 
    xlab("Year") + 
    ylab("") + 
    ggtitle("Cumultative Article Count") +
     scale_x_continuous(breaks=seq(1989, 2015, 8)) +
        scale_y_continuous(breaks = c(seq(0, 35000, 5000)), limits = c(0, 35000))

@@

plot.journal.count <- ggplot(n.journals.year, aes(record.y, 
    y = Journal.count, group=1)) + 
    geom_line(colour=cp[4], size = 1) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=8), 
    title = element_text(size = 10), plot.title=element_text(size=10)) + 
    xlab("Year") + 
    ylab("N") + 
    ggtitle("Journal Count") +
    scale_y_continuous(breaks = c(seq(0, 100, 5)), limits = c(20, 70)) +
     scale_x_discrete(breaks=seq(1989, 2015, 8))

@@

n_authors2.melted <- reshape2::melt(n_authors2, id.vars = c("year", "Standard_Deviation"))
n_authors3.melted <- filter(n_authors2.melted, variable == "Average")

plot.author.count <- ggplot(n_authors2.melted, aes(x = year, y = value, group=variable, colour = variable)) + 
    geom_line() +
    geom_point(data = n_authors3.melted, aes(size = Standard_Deviation), colour = "#e66101") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1), 
    title = element_text(size = 10)) +  
    scale_x_discrete(breaks=c(seq(1989, 2013, 2))) +
    scale_y_continuous(limits = c(.75,2.5)) +
    ylab("Authors") +
    xlab("Year") +
    ggtitle("Number of Authors") +
    labs(size = "Standard deviation") +
    labs(colour = "Value") + 
    scale_colour_manual(values = c("#fdb864", "#e66101"))

@@

df1.summarized <- df1 %>% group_by(journal) %>% 
    summarize(EBP.counts = sum(N)) %>% 
    arrange(desc(EBP.counts)) %>%
    slice(1:15)

journal.evidence.plot <- ggplot(na.omit(df1[df1$journal %in% df1.summarized$journal,]), aes(x = N, y= journal, group = journal, colour = studyType)) + 
    geom_segment(aes(yend = journal), xend=0, colour="grey50") +
    geom_point(size = 2.5) + 
    scale_color_manual(values = cp) + 
    theme_bw() +
    facet_grid(.~studyType) + 
    theme(panel.grid.major.y = element_blank(), legend.position = "none") + 
    xlim(0, 600) + 
    ylab("Journal") +  
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    scale_x_continuous(breaks = seq(0, 60, 10))

@@

df2 <- df2 %>% arrange(desc(Prop)) %>% slice(1:15)

@@

pages.df.melted <- reshape2::melt(pages.df, id.vars = c("years", "Std.Pages"))
pages.df2.melted <- filter(pages.df.melted, variable == "Mean.Pages")

page.plot <- ggplot(pages.df.melted, aes(x = years, y = value, group=variable, colour = variable)) + 
    geom_line() +
    geom_point(data = pages.df2.melted, aes(size = Std.Pages), colour = "#5e3c99") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1), 
    title = element_text(size = 10)) +  
    scale_x_discrete(breaks=c(seq(1989, 2013, 2))) +
    ylab("Pages") +
    xlab("Year") +
    ggtitle("Article Length") +
    labs(size = "Standard deviation") +
    labs(colour = "Value") + 
    scale_colour_manual(values = c("#b2abd2", "#5e3c99"))

@@

# Results
```{r}
# As shown in Table 1, `r nrow(final.list.25)` journals were included in the current study.  The mean number of years of inclusion is `r options(digits=2); mean(final.list.updated$study.time)` (median = `r options(digits=2); median(final.list.updated$study.time);`, sd =  `r options(digits=2); sd(final.list.updated$study.time)`). The overall h-Index for each journal.  The h-Index values represent the current h-Index of each journal, as opposed to the h-Index of the final study period, so these values should be interpreted cautiously.  The mean h-Index value is XXXX, (median = XXX; sd = XXX). A total of `r pretty(sum(articles.year.df$year.count))` article records were obtained across these journals.  The average number of articles per journal was `r options(digits=2); pretty(mean(articles.year.df$year.count))` (median = `r options(digits=2); pretty(median(articles.year.df$year.count))`, sd = `r options(digits=2); pretty(sd(articles.year.df$year.count))`). This figure shows consistent growth in the number of articles from 1989 to 2005.  Notable fluctuations in the number of articles was observed from 2005 to 2008, followed by growth until 2012. The decline from 2012 to 2013 should be interpreted cautiously, as this may be an artifact due to delays in indexing.  
```

@@

Figure 1 uses a small multiples plot to summarize the number of article records
for each year throughout the study period.  According to Tufte (xxxx), small multiples (also referred to as
trellis and lattice charts) are the best design solution for visually enforcing
comparisons of change.  In the current useage, each journal is summarized with its own line plot using two different lines to represent the number of article records contained in the SWHD (orange line) and Scopus (purple line).  The journals are sorted in descending order based on the overall number of article records in the SWHD.  Journal titles are also abbreviated, and the full spellings are available in Appendix A.  These line plots help show publication patterns for each journal.

The differences in publication counts are explained by a couple different reasons.  Scopus does not index journals prior to 1996, whereas the SWHD goes back to 1989.  Although the timeframe of the SWHD is longer than Scopus, the actual number of article records in the SWHD for many journals is lower than what is available in Scopus.  Scopus included editorials and book reviews in their article count, whereas these publications were specifically excluded from the SWHD.  Because Scopus data were available only in aggregate form, it was not possible to extract these data to make more refined comparisons.  In the construction of the SWHD, we also applied a series of cleaning procedures that further reduced the article count for many journals.  For example, we excluded article records that did not contain author names, an abstract or article title.  Again, because we were working only with aggregate data from Scopus, we were unable to apply the same cleaning procedures to the Scopus data.  And, finally, differences can also be attributed to gaps and errors in indexing in Scopus and all the databases used to construct the SWHD.  This is a known and longstanding problem with some social work journals (see Holden, Holden).

@@

@@

```{r}
grid.arrange(plot.journal.count, plot.article.count, ncol=2)

@@

```{r}
stacked.plot
```
\newpage
```{r}
journal.evidence.plot
```

\newpage
```{r}
EBP.plot
```
